---
title: "HW 7 Variance Explained with Two Predictors"
author: "Debi Gonzalez"
date: "2026-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading the dataset
am_data <- read.csv("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/HW 6 Multiple Regression with Two Predictors in R/academicMotivation.csv")
```

For more information on the variables in this dataset, see the accompanying description in the Word Doc in the same folder as this homework assignment and dataset.

1.	Run a regression predicting academic motivation (y) from average hours worked during the week (x1) and advisor supportiveness (x2). Interpret the model R-squared (or adjusted R-squared–your choice) and F test from this model in words.

```{r}
model <- lm(academicMotivation ~ hrs + advisorSupport, data = am_data)
summary(model)
```
The model had a Multiple R-squared of 0.2228 and an Adjusted R-squared of 0.2165. Approximately 22.3% of the variance in academic motivation is explained jointly by average hours worked per week and advisor supportiveness. After adjusting for the number of predictors in the model, about 21.7% of the variance in academic motivation is explained.

The overall regression model was statistically significant with F(2, 247) = 35.40, p < .001. The F-test indicates that, taken together, hours worked and advisor supportiveness significantly predict academic motivation.

2.	Is this a small, moderate, or large R-squared effect size by Cohen’s (1988) rules of thumb?

Cohen (1988) guidelines for R² effect sizes in multiple regression are Small ≈ 0.02, Medium ≈ 0.13, and Large ≈ 0.26. Given our model's R² effect of 0.22, this would be a moderate-to-large effect size according to Cohen’s (1988) guidelines.

3.	How much variance does advisor supportiveness (x2) uniquely explain in academic motivation (y), above and beyond that explained by average hours worked during the week (x1)? Calculate ΔR^2 due to advisor supportiveness (x2) using the methods shown in class. Interpret ΔR^2 and its associated F change significance test in words.

ΔR^2=( t^2 (1-R_full^2))/(df_residual)

•  t=6.070(for advisorSupport)
•  df_residual=247
•  R_full^2=.2228

```{r}
6.070^2

(36.8449*(1-.2228))/247
```
ΔR^2≈.116, which means that advisor supportiveness uniquely explains about 11.6% of the variance in academic motivation, above and beyond hours worked.

F_change=t^2=36.85 --> When adding only one predictor, the F-test for change in R² is mathematically identical to squaring that predictor’s t-value. Adding advisor supportiveness significantly improves prediction of academic motivation beyond hours worked alone, F(1, 247) = 36.85, p < .001.



4.	Is this a small, moderate, or large R-squared effect size by Cohen’s (1988) rules of thumb?

At 0.116, this would be a small-to-moderate effect size according Cohen's rules.
