---
title: "KFF Preliminary Analysis"
author: "Debi Gonzalez"
date: "2026-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # read in sav files
library(tidyverse) # for clean coding
library(lavaan) # for latent variables
library(ggplot2) # for graphs and tables
library(dplyr) # for my flexibility variables
```

```{r}
# Data cleaning and Prep

# Load Datasets
kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Data for JAGS (2.19.26).sav")
weekly_kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Weekly Therapist Session Summary data (LONG) abridged for grad students 3.7.26.sav")

# my two variables of interest: Sess_Fit and Sess_Tail
# Both of which constitute our key flexibility variables. These are in long format.


# Flexibility Variables: creating a sub-dataset of the weekly dataset averaging flexibility measures to then join to our main dataset
flexibility_avg_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    avg_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else mean(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    avg_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else mean(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )


```


First, it characterizes the scope of fit and tailoring across sessions by defining case-level indicators of average fit, average tailoring, and, secondarily, variability in fit and tailoring across treatment. 

-- Identify mean and standard deviation of fit and tailoring across each case in the sample. This creates new variables, where each case/child has a mean and sd.

```{r}
# this will use kff_data weekly to create the mean and sd for each individual child

# This can then be joined with the kff_data sub-dataset

```


-- Identify mean and standard deviation of fit and tailoring across sample.
```{r}
# This would go with the summary statistics. 
```


Second, it tests whether baseline child clinical characteristics, caregiver clinical complexity, and demographic factors predict these fit and tailoring indices. 

Model 1: Predicting Mean Fit
MeanFit_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i

Model 2: Predicting Mean Tailoring
MeanTailoring_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i


Third, it tests whether average fit and average tailoring predict anxiety response outcome and treatment satisfaction (for caregiver and child). 

Model 3: Predicting Anxiety Outcome
PostAnxiety_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 4: Predicting Caregiver Treatment Satisfaction
CaregiverSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 5: Predicting Youth Treatment Satisfaction
YouthSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i