---
title: "KFF Preliminary Analysis"
author: "Debi Gonzalez"
date: "2026-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # read in sav files
library(tidyverse) # for clean coding
library(lavaan) # for latent variables
library(ggplot2) # for graphs and tables
library(dplyr) # for my flexibility variables
library(gt) # for formatting my beautiful tables :)
```

```{r}
# Data cleaning and Prep

# Load Datasets
kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Data for JAGS (2.19.26).sav")
weekly_kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Weekly Therapist Session Summary data (LONG) abridged for grad students 3.7.26.sav")

# my two variables of interest: Sess_Fit and Sess_Tail
# Both of which constitute our key flexibility variables. These are in long format.


# Flexibility Variables: creating a sub-dataset of the weekly dataset averaging flexibility measures to then join to our main dataset
flexibility_avg_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    avg_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else mean(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    avg_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else mean(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

# Doing the same for SD variables
flexibility_sd_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    sd_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else sd(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    sd_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else sd(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

# Identifying the therapist-led CBT sub-sample
cbt_subsample <- kff_data %>%
  filter(TxGroup_0 == 0) %>% # treatment group = 0 is the therapist led sub-sample
  select(
    ID,
    # Baseline anxiety
    PROMt_C_B, PROMt_Y_B, PARS_Tot_B,
    # Behavioral problems (externalizing and anxiety)
    PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B,
    # Caregiver complexity (accomodation and clinical complexity)
    PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B,
    # ACEs
    ACEQ_Tot_pool_B, ACEQ4_pool_B,
    # Caregiver education
    Educ3_C,
    # Post-treatment anxiety
    PROMt_C_P, PROMt_Y_P, PARS_Tot_P,
    # Post-treatment caregiver satisfaction
    SATQTot_C_P, SATQ_Mean_C_P_0,
    # Post-treatment youth satisfaction
    SATQtot_Y_P, SATQ_Mean_Y_P_0,
    # Main outcomes control variables
    InsecurityStatus_Pool, Male_Y_pool, AgeGrp_0, Lang_0, Black_Y_pool, Hisp_Y_pool, Site1, Site2, Site3, Site4)
  
# Notes to self on variable selection  
  # Variables needed for latent baseline anxiety: PROMt_C_B, PROMt_Y_B, PARS_Tot_B
  # Variables needed for behavioral problems: PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B
  # Variables needed for caregiver complexity: PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B
  # variables needed for ACEs: ACEQ_Tot_pool_B, ACEQ4_pool_B
  # Variables needed for Caregiver Education: Educ3_C
  # Variables needed for Post-treatment Anxiety: PROMt_C_P, PROMt_Y_P, PARS_Tot_P
  # Variables needed for Post-Treatment Caregiver Satisfaction: SATQTot_C_P, SATQ_Mean_C_P_0
  # Variables needed for Post-Treatment Youth Satisfaction: SATQTot_Y_P, SATQ_Mean_Y_P_0
  # Main Outcomes control variables: InsecurityStatus_Pool, Male_Y_Pool, AgeGrp0, Lang_0, Black_Y_pool, Hisp_Y_pool
  
  
  
# Per Main Outcomes paper: All models included the following covariates: Region/Site, Language of Care, Youth Age Group, Youth Gender, Black Race, Hispanic Ethnicity, and Household Resource Insecurity.   
  
#Checking what my variable numbers mean
  attr(kff_data$Educ3_C, "labels")
# Caregiver Education (Highest Level Completed) 1 = High School or Less, 2 = Completed College, 3 = Completed Graduate School, No response
  
  
  
# Now joining the average flexibility scores to the cbt_subsample

# left_join test  
test1 <- cbt_subsample %>%
  left_join(flexibility_avg_kff_data, by = "ID")

# inner_join test
test2 <- cbt_subsample %>%
  inner_join(flexibility_avg_kff_data, by = "ID") # this gives us a joint dataset with only 145 cases, so I'm going to stick to the left join that has my 149 CBT cases regardless of whether we have flexibility score.

# The practical difference: left_join() keeps all 149 therapist-led CBT rows, even if some have no flexibility match while inner_join() keeps only rows where the ID exists in both datasets


cbt_kff_joined <- cbt_subsample %>% # this subsample isolates the 149 therapist led CBT cases
  left_join(flexibility_avg_kff_data, by = "ID") %>% # joining the avg flexibility by child
  left_join(flexibility_sd_kff_data, by = "ID") # joining the sd flexibility by child

```


```{r}
# Building the latent variables: baseline anxiety severity, baseline behavioral problems, baseline caregiver complexity, post-treatment anxiety severity

# library lavaan is already loaded --> best practice for latent variables bc the latent factor is estimated from the covariance structure, so indicators can be on different scales.

latent_model1 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems =~ PSC_Ext_Y_B + PSC_Ext_C_B + PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ PABUover_C_B + DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'

fit1 <- cfa(
  latent_model1,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit1, fit.measures = TRUE, standardized = TRUE)


# Based on these findings, I have decided to create a new set of latent variables that may be more appropriate.

latent_model2 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems =~ PSC_Ext_Y_B + PSC_Ext_C_B + PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'
# The main change is removing caregiver accomodation measure from the caregiver complexity latent variable as it barely loads on the factor and it's positive while the other three DASS variables are negative.

fit2 <- cfa(
  latent_model2,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit2, fit.measures = TRUE, standardized = TRUE)

# Removing PABUover_C_B helped:
  # the caregiver factor no longer has mixed signs
  # the three DASS indicators now all load strongly and positively
  # SRMR improved from 0.077 to 0.069
  # CFI improved from 0.823 to 0.862
  # TLI improved from 0.773 to 0.818
  # RMSEA improved a little from 0.090 to 0.085


# Let's try one more time creating two constructs of behavior problems: one for attention and one for externalizing problems.

latent_model3 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems_ext =~ PSC_Ext_Y_B + PSC_Ext_C_B
  behavior_problems_att =~ PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'
# The main change is removing caregiver accomodation measure from the caregiver complexity latent variable as it barely loads on the factor and it's positive while the other three DASS variables are negative.

fit3 <- cfa(
  latent_model3,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit3, fit.measures = TRUE, standardized = TRUE)

modificationIndices(fit3, sort. = TRUE, minimum.value = 10)
# This says the model sees only one main leftover source of misfit: it suggests that DASSs_C_B might also load on behavior_problems_ext. Conceptually though I do not want to add caregiver stress to the measure of child externalizing problems, even if they are correlated. Caregiver stress (DASSs_C_B) may share variance with externalizing problems that is not fully captured by the current factor structure.

# A revised CFA separating behavioral problems into externalizing and attention factors showed improved fit relative to prior models, χ²(55) = 102.08, p < .001, CFI = .898, TLI = .855, RMSEA = .076, SRMR = .062. Standardized loadings were moderate to strong for baseline anxiety, caregiver complexity, and post-treatment anxiety, and the split behavioral factors showed significant loadings for both externalizing and attention indicators. The externalizing and attention factors were moderately correlated, supporting their related but distinct nature.


# Adding the latent variables to the main analysis dataset:

fscores <- as.data.frame(lavPredict(fit3)) # creates the latent score columns in data frame format
idx <- lavInspect(fit3, "case.idx") # gives the exact row positions those scores belong to

cbt_kff_joined_latent <- cbt_kff_joined # Creating a new dataset that includes the latent variables
cbt_kff_joined_latent[colnames(fscores)] <- NA_real_ # creates all needed latent columns at once
cbt_kff_joined_latent[idx, colnames(fscores)] <- fscores # this fills in the correct rows

# Checking that my data combined correctly:
length(idx) == nrow(fscores)
cbt_kff_joined[idx, "ID"]
cbt_kff_joined_latent[idx, "ID"]


# Note on latent variable scale interpretation: they are factor scores estimated by lavPredict(), not observed raw scores. Each latent variable was scaled by fixing its variance to 1 in the model, instead of fixing the first loading to 1. Latent means are fixed to 0 by default in a single-group model.So, A score like 0 means the person is around the sample’s latent mean, +1 means about one latent standard deviation above the sample mean, -1 means about one latent standard deviation below the sample mean. This means that higher scores indicate more of the latent construct, scores are centered approximately around zero, and a 1-unit difference is roughly a 1-SD difference on the latent factor scale.


# Now that I have my latent variables, I'm going to add some more demographic variables from the original dataset for a descriptive statistics table

race_gender_vars <- kff_data %>%
  select(
    ID,
    Gender_Y_pool,
    White_Y_pool,
    AmInd_Y_pool,
    Asian_Y_pool,
    NatHPI_Y_pool,
    NotList_Y_pool,
    NonHispWhite_Y_pool)

cbt_kff_joined_latent <- cbt_kff_joined_latent %>%
  left_join(race_gender_vars, by = "ID")

# For this demographics table, given my sample size, I'm creating a summary race variable:

cbt_kff_joined_latent <- cbt_kff_joined_latent %>%
  mutate(
    Other_Racial_Groups = if_else(
      AmInd_Y_pool == 1 |
      Asian_Y_pool == 1 |
      NatHPI_Y_pool == 1 |
      NotList_Y_pool == 1,
      1,0)) # Missing Middle Eastern/Northern African bc it was not included in the raw dataset provided by the lab
# Checking my numbers
cbt_kff_joined_latent %>%
  count(Other_Racial_Groups) %>% 
  view()
# Note that this does not 100% match the main outcomes paper bc of the missing middle eastern and nothern african variable.

# Extra verification of my numbers.
kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(NotList_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(AmInd_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(Asian_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(NatHPI_Y_pool)


```

First, it characterizes the scope of fit and tailoring across sessions by defining case-level indicators of average fit, average tailoring, and, secondarily, variability in fit and tailoring across treatment. 


```{r}

# Demographics Summary Table (Age group, sex, race/ethnicity, resource instability, parental education)


# Flexibility across the sample: a table of summary statistics


# Let's visualize the distribution of mean fit and tailoring across the sample --> histogram


```

```{r}
# Initial table of *descriptive statistics* with means, SDs, and lower triangle of the correlation matrix for the following variables: fit, tailoring, baseline anxiety severity, ACEs Total score, externalizing behavior problems, attention behavior problems, caregiver complexity, caregiver accommodation, post-treatment anxiety severity, caregiver treatment satisfaction, youth treatment satisfaction
```
From my cbt_kff_joined_latent dataset, I want to create a table of descriptive statistics with means, SDs, and lower triangle of the correlation matrix of

Second, it tests whether baseline child clinical characteristics, caregiver clinical complexity, and demographic factors predict these fit and tailoring indices. 

Model 1: Predicting Mean Fit
MeanFit_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i

Model 2: Predicting Mean Tailoring
MeanTailoring_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i


Third, it tests whether average fit and average tailoring predict anxiety response outcome and treatment satisfaction (for caregiver and child). 

Model 3: Predicting Anxiety Outcome
PostAnxiety_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 4: Predicting Caregiver Treatment Satisfaction
CaregiverSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 5: Predicting Youth Treatment Satisfaction
YouthSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i