---
title: "KFF Preliminary Analysis"
author: "Debi Gonzalez"
date: "2026-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # read in sav files
library(tidyverse) # for clean coding
library(lavaan) # for latent variables
library(ggplot2) # for graphs and tables
library(dplyr) # for my flexibility variables
```

```{r}
# Data cleaning and Prep

# Load Datasets
kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Data for JAGS (2.19.26).sav")
weekly_kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Weekly Therapist Session Summary data (LONG) abridged for grad students 3.7.26.sav")

# my two variables of interest: Sess_Fit and Sess_Tail
# Both of which constitute our key flexibility variables. These are in long format.


# Flexibility Variables: creating a sub-dataset of the weekly dataset averaging flexibility measures to then join to our main dataset
flexibility_avg_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    avg_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else mean(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    avg_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else mean(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

flexibility_sd_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    sd_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else sd(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    sd_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else sd(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

# Identifying the therapist-led CBT sub-sample
cbt_subsample <- kff_data %>%
  filter(TxGroup_0 == 0) %>% # treatment group = 0 is the therapist led sub-sample
  select(
    ID,
    # Baseline anxiety
    PROMt_C_B, PROMt_Y_B, PARS_Tot_B,
    # Behavioral problems
    PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B,
    # Caregiver complexity
    PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B,
    # ACEs
    ACEQ_Tot_pool_B, ACEQ4_pool_B,
    # Caregiver education
    Educ3_C,
    # Post-treatment anxiety
    PROMt_C_P, PROMt_Y_P, PARS_Tot_P,
    # Post-treatment caregiver satisfaction
    SATQTot_C_P, SATQ_Mean_C_P_0,
    # Post-treatment youth satisfaction
    SATQtot_Y_P, SATQ_Mean_Y_P_0,
    # Main outcomes control variables
    InsecurityStatus_Pool, Male_Y_pool, AgeGrp_0, Lang_0, Black_Y_pool, Hisp_Y_pool, Site1, Site2, Site3, Site4)
  
# Notes to self on variable selection  
  # Variables needed for latent baseline anxiety: PROMt_C_B, PROMt_Y_B, PARS_Tot_B
  # Variables needed for behavioral problems: PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B
  # Variables needed for caregiver complexity: PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B
  # variables needed for ACEs: ACEQ_Tot_pool_B, ACEQ4_pool_B
  # Variables needed for Caregiver Education: Educ3_C
  # Variables needed for Post-treatment Anxiety: PROMt_C_P, PROMt_Y_P, PARS_Tot_P
  # Variables needed for Post-Treatment Caregiver Satisfaction: SATQTot_C_P, SATQ_Mean_C_P_0
  # Variables needed for Post-Treatment Youth Satisfaction: SATQTot_Y_P, SATQ_Mean_Y_P_0
  # Main Outcomes control variables: InsecurityStatus_Pool, Male_Y_Pool, AgeGrp0, Lang_0, Black_Y_pool, Hisp_Y_pool
  
  
  
# Per Main Outcomes paper: All models included the following covariates: Region/Site, Language of Care, Youth Age Group, Youth Gender, Black Race, Hispanic Ethnicity, and Household Resource Insecurity.   
  
#Checking what my variable numbers mean
  attr(kff_data$Educ3_C, "labels")
# Caregiver Education (Highest Level Completed) 1 = High School or Less, 2 = Completed College, 3 = Completed Graduate School, No response
  
  
  
# Now joining the average flexibility scores to the cbt_subsample

# left_join test  
test1 <- cbt_subsample %>%
  left_join(flexibility_avg_kff_data, by = "ID")

# inner_join test
test2 <- cbt_subsample %>%
  inner_join(flexibility_avg_kff_data, by = "ID") # this gives us a joint dataset with only 145 cases, so I'm going to stick to the left join that has my 149 CBT cases regardless of whether we have flexibility score.

# The practical difference: left_join() keeps all 149 therapist-led CBT rows, even if some have no flexibility match while inner_join() keeps only rows where the ID exists in both datasets


cbt_kff_joined <- cbt_subsample %>% # this subsample isolates the 149 therapist led CBT cases
  left_join(flexibility_avg_kff_data, by = "ID") %>% # joining the avg flexibility by child
  left_join(flexibility_sd_kff_data, by = "ID") # joining the sd flexibility by child

```


```{r}
# Building the latent variables


```

First, it characterizes the scope of fit and tailoring across sessions by defining case-level indicators of average fit, average tailoring, and, secondarily, variability in fit and tailoring across treatment. 

-- Identify mean and standard deviation of fit and tailoring across each case in the sample. This creates new variables, where each case/child has a mean and sd.

```{r}
# this will use kff_data weekly to create the mean and sd for each individual child

# This can then be joined with the kff_data sub-dataset

```


-- Identify mean and standard deviation of fit and tailoring across sample.
```{r}
# This would go with the summary statistics. 
```


Second, it tests whether baseline child clinical characteristics, caregiver clinical complexity, and demographic factors predict these fit and tailoring indices. 

Model 1: Predicting Mean Fit
MeanFit_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i

Model 2: Predicting Mean Tailoring
MeanTailoring_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i


Third, it tests whether average fit and average tailoring predict anxiety response outcome and treatment satisfaction (for caregiver and child). 

Model 3: Predicting Anxiety Outcome
PostAnxiety_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 4: Predicting Caregiver Treatment Satisfaction
CaregiverSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

Model 5: Predicting Youth Treatment Satisfaction
YouthSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i