---
title: "KFF Preliminary Analysis"
author: "Debi Gonzalez"
date: "2026-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(haven) # read in sav files
library(tidyverse) # for clean coding
library(tidyr) # for pivot_longer and other table manipulations and data cleaning
library(lavaan) # for latent variables
library(ggplot2) # for graphs and tables
library(dplyr) # for my flexibility variables and later tables (data wrangling)
library(gt) # for formatting my beautiful tables :)
library(tibble) # another library for tables that lets us manually build small data frames
library(broom) # to transform model output into tidy data frames (i.e. extract the regression coefficients, standard errors, t values, and p values)
```

```{r}
# Data cleaning and Prep

# Load Datasets
kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Data for JAGS (2.19.26).sav")
weekly_kff_data <- read_sav("/Users/debig/Documents/PhD Y1/Stats II/Quant II Projects/KFF_Quant_II/KFF Weekly Therapist Session Summary data (LONG) abridged for grad students 3.7.26.sav")

# my two variables of interest: Sess_Fit and Sess_Tail
# Both of which constitute our key flexibility variables. These are in long format.


# Flexibility Variables: creating a sub-dataset of the weekly dataset averaging flexibility measures to then join to our main dataset
flexibility_avg_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    avg_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else mean(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    avg_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else mean(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

# Doing the same for SD variables
flexibility_sd_kff_data <- weekly_kff_data %>%
  group_by(ID) %>%
  summarise(
    # if all NAs for one ID, then avg variable will be NA
    sd_Sess_Fit = if (all(is.na(Sess_Fit))) NA_real_ else sd(Sess_Fit, na.rm = TRUE), # na.rm = TRUE is telling R to ignore NA values for now
    sd_Sess_Tail = if (all(is.na(Sess_Tail))) NA_real_ else sd(Sess_Tail, na.rm = TRUE),
    .groups = "drop" # ungrouping for future ease
  )

# Identifying the therapist-led CBT sub-sample
cbt_subsample <- kff_data %>%
  filter(TxGroup_0 == 0) %>% # treatment group = 0 is the therapist led sub-sample
  select(
    ID,
    # Baseline anxiety
    PROMt_C_B, PROMt_Y_B, PARS_Tot_B,
    # Behavioral problems (externalizing and anxiety)
    PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B,
    # Caregiver complexity (accomodation and clinical complexity)
    PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B,
    # ACEs
    ACEQ_Tot_pool_B, ACEQ4_pool_B,
    # Caregiver education
    Educ3_C,
    # Post-treatment anxiety
    PROMt_C_P, PROMt_Y_P, PARS_Tot_P,
    # Post-treatment caregiver satisfaction
    SATQTot_C_P, SATQ_Mean_C_P_0,
    # Post-treatment youth satisfaction
    SATQtot_Y_P, SATQ_Mean_Y_P_0,
    # Main outcomes control variables
    InsecurityStatus_Pool, Male_Y_pool, AgeGrp_0, Lang_0, Black_Y_pool, Hisp_Y_pool, Site1, Site2, Site3, Site4)
  
# Notes to self on variable selection  
  # Variables needed for latent baseline anxiety: PROMt_C_B, PROMt_Y_B, PARS_Tot_B
  # Variables needed for behavioral problems: PSC_Ext_Y_B, PSC_Ext_C_B, PSC_Att_C_B, PSC_Att_Y_B
  # Variables needed for caregiver complexity: PABUover_C_B, DASSd_C_B, DASSa_C_B, DASSs_C_B
  # variables needed for ACEs: ACEQ_Tot_pool_B, ACEQ4_pool_B
  # Variables needed for Caregiver Education: Educ3_C
  # Variables needed for Post-treatment Anxiety: PROMt_C_P, PROMt_Y_P, PARS_Tot_P
  # Variables needed for Post-Treatment Caregiver Satisfaction: SATQTot_C_P, SATQ_Mean_C_P_0
  # Variables needed for Post-Treatment Youth Satisfaction: SATQTot_Y_P, SATQ_Mean_Y_P_0
  # Main Outcomes control variables: InsecurityStatus_Pool, Male_Y_Pool, AgeGrp0, Lang_0, Black_Y_pool, Hisp_Y_pool
  
  
  
# Per Main Outcomes paper: All models included the following covariates: Region/Site, Language of Care, Youth Age Group, Youth Gender, Black Race, Hispanic Ethnicity, and Household Resource Insecurity.   
  
#Checking what my variable numbers mean
  attr(kff_data$Educ3_C, "labels")
# Caregiver Education (Highest Level Completed) 1 = High School or Less, 2 = Completed College, 3 = Completed Graduate School, No response
  
  
  
# Now joining the average flexibility scores to the cbt_subsample

# left_join test  
test1 <- cbt_subsample %>%
  left_join(flexibility_avg_kff_data, by = "ID")

# inner_join test
test2 <- cbt_subsample %>%
  inner_join(flexibility_avg_kff_data, by = "ID") # this gives us a joint dataset with only 145 cases, so I'm going to stick to the left join that has my 149 CBT cases regardless of whether we have flexibility score.

# The practical difference: left_join() keeps all 149 therapist-led CBT rows, even if some have no flexibility match while inner_join() keeps only rows where the ID exists in both datasets


cbt_kff_joined <- cbt_subsample %>% # this subsample isolates the 149 therapist led CBT cases
  left_join(flexibility_avg_kff_data, by = "ID") %>% # joining the avg flexibility by child
  left_join(flexibility_sd_kff_data, by = "ID") # joining the sd flexibility by child

```


```{r}
# Building the latent variables: baseline anxiety severity, baseline behavioral problems, baseline caregiver complexity, post-treatment anxiety severity

# library lavaan is already loaded --> best practice for latent variables bc the latent factor is estimated from the covariance structure, so indicators can be on different scales.

latent_model1 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems =~ PSC_Ext_Y_B + PSC_Ext_C_B + PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ PABUover_C_B + DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'

fit1 <- cfa(
  latent_model1,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit1, fit.measures = TRUE, standardized = TRUE)


# Based on these findings, I have decided to create a new set of latent variables that may be more appropriate.

latent_model2 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems =~ PSC_Ext_Y_B + PSC_Ext_C_B + PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'
# The main change is removing caregiver accomodation measure from the caregiver complexity latent variable as it barely loads on the factor and it's positive while the other three DASS variables are negative.

fit2 <- cfa(
  latent_model2,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit2, fit.measures = TRUE, standardized = TRUE)

# Removing PABUover_C_B helped:
  # the caregiver factor no longer has mixed signs
  # the three DASS indicators now all load strongly and positively
  # SRMR improved from 0.077 to 0.069
  # CFI improved from 0.823 to 0.862
  # TLI improved from 0.773 to 0.818
  # RMSEA improved a little from 0.090 to 0.085


# Let's try one more time creating two constructs of behavior problems: one for attention and one for externalizing problems.

latent_model3 <- 'baseline_anxiety =~ PROMt_C_B + PROMt_Y_B + PARS_Tot_B
  behavior_problems_ext =~ PSC_Ext_Y_B + PSC_Ext_C_B
  behavior_problems_att =~ PSC_Att_C_B + PSC_Att_Y_B
  caregiver_complexity =~ DASSd_C_B + DASSa_C_B + DASSs_C_B
  post_anxiety =~ PROMt_C_P + PROMt_Y_P + PARS_Tot_P'
# The main change is removing caregiver accomodation measure from the caregiver complexity latent variable as it barely loads on the factor and it's positive while the other three DASS variables are negative.

fit3 <- cfa(
  latent_model3,
  data = cbt_kff_joined,
  missing = "fiml", # telling my model to use full-information maximum likelihood to handle missing data, instead of the default listwise deletion; it is not creating imputed values you can inspect in the dataset
  std.lv = TRUE) # telling lavaan to fix each latent variable’s variance to 1 so the factor is put on a standardized scale.

summary(fit3, fit.measures = TRUE, standardized = TRUE)

modificationIndices(fit3, sort. = TRUE, minimum.value = 10)
# This says the model sees only one main leftover source of misfit: it suggests that DASSs_C_B might also load on behavior_problems_ext. Conceptually though I do not want to add caregiver stress to the measure of child externalizing problems, even if they are correlated. Caregiver stress (DASSs_C_B) may share variance with externalizing problems that is not fully captured by the current factor structure.

# A revised CFA separating behavioral problems into externalizing and attention factors showed improved fit relative to prior models, χ²(55) = 102.08, p < .001, CFI = .898, TLI = .855, RMSEA = .076, SRMR = .062. Standardized loadings were moderate to strong for baseline anxiety, caregiver complexity, and post-treatment anxiety, and the split behavioral factors showed significant loadings for both externalizing and attention indicators. The externalizing and attention factors were moderately correlated, supporting their related but distinct nature.


# Adding the latent variables to the main analysis dataset:

fscores <- as.data.frame(lavPredict(fit3)) # creates the latent score columns in data frame format
idx <- lavInspect(fit3, "case.idx") # gives the exact row positions those scores belong to

cbt_kff_joined_latent <- cbt_kff_joined # Creating a new dataset that includes the latent variables
cbt_kff_joined_latent[colnames(fscores)] <- NA_real_ # creates all needed latent columns at once
cbt_kff_joined_latent[idx, colnames(fscores)] <- fscores # this fills in the correct rows

# Checking that my data combined correctly:
length(idx) == nrow(fscores)
cbt_kff_joined[idx, "ID"]
cbt_kff_joined_latent[idx, "ID"]


# Note on latent variable scale interpretation: they are factor scores estimated by lavPredict(), not observed raw scores. Each latent variable was scaled by fixing its variance to 1 in the model, instead of fixing the first loading to 1. Latent means are fixed to 0 by default in a single-group model.So, A score like 0 means the person is around the sample’s latent mean, +1 means about one latent standard deviation above the sample mean, -1 means about one latent standard deviation below the sample mean. This means that higher scores indicate more of the latent construct, scores are centered approximately around zero, and a 1-unit difference is roughly a 1-SD difference on the latent factor scale.


# Now that I have my latent variables, I'm going to add some more demographic variables from the original dataset for a descriptive statistics table

race_gender_vars <- kff_data %>%
  select(
    ID,
    Gender_Y_pool,
    White_Y_pool,
    AmInd_Y_pool,
    Asian_Y_pool,
    NatHPI_Y_pool,
    NotList_Y_pool,
    NonHispWhite_Y_pool)

cbt_kff_joined_latent <- cbt_kff_joined_latent %>%
  left_join(race_gender_vars, by = "ID")

# For this demographics table, given my sample size, I'm creating a summary race variable:

cbt_kff_joined_latent <- cbt_kff_joined_latent %>%
  mutate(
    Other_Racial_Groups = if_else(
      AmInd_Y_pool == 1 |
      Asian_Y_pool == 1 |
      NatHPI_Y_pool == 1 |
      NotList_Y_pool == 1,
      1,0)) # Missing Middle Eastern/Northern African bc it was not included in the raw dataset provided by the lab
# Checking my numbers
cbt_kff_joined_latent %>%
  count(Other_Racial_Groups) %>% 
  view()
# Note that this does not 100% match the main outcomes paper bc of the missing middle eastern and nothern african variable.

# Extra verification of my numbers.
kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(NotList_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(AmInd_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(Asian_Y_pool)

kff_data %>% 
  filter(TxGroup_0 == 0) %>% 
  count(NatHPI_Y_pool)


```

First, it characterizes the scope of fit and tailoring across sessions by defining case-level indicators of average fit, average tailoring, and, secondarily, variability in fit and tailoring across treatment. 


```{r}

# Demographics Summary Table (Age group, sex, race/ethnicity, resource instability, parental education)


# Building the main demographics table data. This part creates all the demographic sections EXCEPT race/ethnicity (not mutually exclusive)
demo_table_data <- cbt_kff_joined_latent %>%
  transmute(
    # transmute() creates a new data frame using only the variables listed here. It drops everything else, so we can have a clean dataframe for the table.

    `Caregiver Education` = case_when(
      Educ3_C == 1 ~ "High School or Less",
      Educ3_C == 2 ~ "Completed College",
      Educ3_C == 3 ~ "Completed Graduate School",
      TRUE ~ "No response"),
    # Recoding Educ3_C into readable category labels. Any value not equal to 1, 2, or 3 becomes "No response".

    `Resource Instability` = case_when(
      InsecurityStatus_Pool == 0 ~ "Did not endorse housing/resource insecurity",
      InsecurityStatus_Pool == 1 ~ "Endorsed housing/resource insecurity",
      TRUE ~ "No response"),
    # Recoding binary insecurity variable into readable labels.

    `Sex/Gender` = case_when(
      Gender_Y_pool == 1 ~ "Male",
      Gender_Y_pool == 2 ~ "Female",
      Gender_Y_pool == 3 ~ "Non-binary",
      Gender_Y_pool == 4 ~ "Gender fluid",
      Gender_Y_pool == 5 ~ "Trans male",
      Gender_Y_pool == 6 ~ "Trans female",
      Gender_Y_pool == 7 ~ "Different identity",
      Gender_Y_pool == 888 ~ "Prefer not to answer",
      TRUE ~ "No response"),
    # Recoding gender variable into readable categories.

    `Age Group` = case_when(
      AgeGrp_0 == 0 ~ "7-12 years",
      AgeGrp_0 == 1 ~ "13+ years",
      TRUE ~ "No response"),
    # Recoding age group into readable labels.

    `Language` = case_when(
      Lang_0 == 0 ~ "English",
      Lang_0 == 1 ~ "Spanish",
      TRUE ~ "No response"),
    # Recoding language variable into readable labels.

    `Region` = case_when(
      Site1 == 1 ~ "Baltimore",
      Site2 == 1 ~ "Boston",
      Site3 == 1 ~ "Miami",
      Site4 == 1 ~ "Seattle",
      TRUE ~ "No response") # Recoding site indicators into one readable region label. Each person belongs to one site only.
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Characteristic",
    values_to = "Category") %>%
  # pivot_longer() reshapes the data from wide to long format. Before: one row per person, many demographic columns. After: one row per person per demographic variable. This would be more suitable for summaries.

  count(Characteristic, Category, name = "n") %>%
  # Counting how many times each Category (the labels of the variables) appears within each Characteristic.

  group_by(Characteristic) %>%
  mutate(
    percent = 100 * n / sum(n)) %>%
  # Now calculating the percentage of each characteristic within a variable (i.e. the labels/categories).
  ungroup()
  # Removing grouping so later operations act on the full data frame normally.


# Building the race/ethnicity separately as a multi-response section. Each indicator is counted independently, so percentages do not need to sum to 100%.
# This is because participants were allowed to choose more than 1 category.

race_table_data <- tibble(
  Characteristic = "Race/Ethnicity",
  # Creating a small manual data frame where every row belongs to the "Race/Ethnicity" section of the final table.

  Category = c(
    "Non-Hispanic White",
    "Hispanic",
    "Black",
    "Other Racial Groups",
    "No Response"),
  # These are the row labels that will appear under Race/Ethnicity.

  n = c(
    sum(cbt_kff_joined_latent$NonHispWhite_Y_pool == 1, na.rm = TRUE),
    sum(cbt_kff_joined_latent$Hisp_Y_pool == 1, na.rm = TRUE),
    sum(cbt_kff_joined_latent$Black_Y_pool == 1, na.rm = TRUE),
    sum(cbt_kff_joined_latent$Other_Racial_Groups == 1, na.rm = TRUE),
    sum(is.na(cbt_kff_joined_latent$NotList_Y_pool)))
  # For each indicator, counting how many people have value 1. Counting NAs as no response.
) %>%
  mutate(percent = 100 * n / nrow(cbt_kff_joined_latent))
  # Calculating the race percentages using the FULL sample size as denominator, not the sum of race counts.


# Now, we have to combine the main demographics data with the race data
demo_table_data <- bind_rows(
  demo_table_data,
  race_table_data)
# Stacking the race rows underneath the rows from the main demographic table. Now the demo_table_data contains all necessary demographic sections.

# Specifying which sections should be sorted by descending count; this variable will be used to specify this below.
sort_by_n <- c("Race/Ethnicity", "Region", "Resource Instability", "Sex/Gender")
# These section names will be sorted so the largest n category appears first.

# Applying the ordering rules: Some sections should be sorted by descending n. Others should keep a logical order defined manually.
demo_table_data <- bind_rows(
  demo_table_data %>%
    filter(Characteristic %in% sort_by_n) %>%
    # Keeping only the sections that should be sorted by descending count.

    group_by(Characteristic) %>%
    arrange(desc(n), .by_group = TRUE) %>%
    # Within each selected section, sorting rows from highest n to lowest n.

    mutate(order = row_number()) %>%
    # Creating an order variable: 1, 2, 3, ... This will be used later to preserve the sorted order.

    ungroup(),

  demo_table_data %>%
    filter(!Characteristic %in% sort_by_n) %>%
    # Now, only keeping sections that should NOT be sorted by n.

    mutate(
      order = case_when(
        Characteristic == "Age Group" & Category == "7-12 years" ~ 1,
        Characteristic == "Age Group" & Category == "13+ years" ~ 2,
        Characteristic == "Age Group" & Category == "No response" ~ 3,

        Characteristic == "Language" & Category == "English" ~ 1,
        Characteristic == "Language" & Category == "Spanish" ~ 2,
        Characteristic == "Language" & Category == "No response" ~ 3,

        Characteristic == "Caregiver Education" & Category == "High School or Less" ~ 1,
        Characteristic == "Caregiver Education" & Category == "Completed College" ~ 2,
        Characteristic == "Caregiver Education" & Category == "Completed Graduate School" ~ 3,
        Characteristic == "Caregiver Education" & Category == "No response" ~ 4,

        TRUE ~ 99))
    # Manually assigning a display order for sections where I want a meaningful sequence instead of descending frequency.
) %>%
  mutate(
    Category = paste0("\u00A0\u00A0\u00A0", Category),
    # Adding leading non-breaking spaces to indent category labels under each heading. There might be a better way to do this, but this is what I got now.
    `n (%)` = sprintf("%d (%.1f%%)", n, percent)
    # Formatting count and percentage into a single display column.
  ) %>%
  arrange(
    factor(Characteristic,
      levels = c(
        "Age Group",
        "Sex/Gender",
        "Race/Ethnicity",
        "Language",
        "Region",
        "Caregiver Education",        
        "Resource Instability")),
    order)
  # Arranging the table rows so the demographic sections appear in a cohesive order and within each section we will be using the previously defined row order.



# Turning the prepared data into a gt table
demo_gt <- demo_table_data %>%
  select(Characteristic, Category, `n (%)`) %>%
  # Keeping only the columns needed for the final display table.

  gt(groupname_col = "Characteristic") %>%
  # Creating a gt table and using Characteristic as the row-group heading. This makes Age Group, Sex/Gender, etc. appear as grouped section headers.
  cols_label(
    Category = "",
    `n (%)` = "n (%)") %>%
  # Setting the displayed column labels. Category column gets a blank header for a cleaner look.
  tab_header(
    title = md("**Therapist-Led CBT Sample Demographics**")) %>%
  # Adding a table title and bolding it.
  tab_source_note(source_note = md("*Note.* Participants could select multiple race/ethnicity categories, so race/ethnicity categories were counted independently. Percentages in this section may not sum to 100%.")) %>%
  # Adding the explanatory note under the table for the race/ethnicity variable design.
  opt_table_font(font = list("Times New Roman", "Times")) %>%
  # Setting two font option Times New Roman first, if not available trying Times.
  tab_style(style = cell_text(weight = "bold"),
    locations = cells_row_groups()) %>%
  # Making each section heading bold.
#  tab_style(style = cell_text(weight = "bold"),
#    locations = cells_title(groups = "title")) %>%
  # Making the table title bold, but not necessary since I used the markdown syntac in the tab_header prompt.
  cols_align(align = "left", columns = Category) %>%
  # Left-aligning the category text.
  cols_align(align = "center", columns = `n (%)`) %>%
  # Centering the n (%) column.
  cols_width(
    Category ~ px(260),
    `n (%)` ~ px(90)
  ) %>%
  tab_options(table.font.size = px(14),
    heading.title.font.size = px(16),
    row_group.font.size = px(14),
    data_row.padding = px(4),
    row_group.padding = px(5),
    table.border.top.width = px(1.5),
    table.border.bottom.width = px(1.5),
    heading.border.bottom.width = px(1),
    row_group.border.top.width = px(1),
    table_body.border.bottom.width = px(1))
  # Adjusting font sizes, row spacing, and border widths to make the table look cleaner.

# Printing the final formatted table
demo_gt
```


```{r}
# Let's visualize the distribution of mean fit and tailoring across the sample --> histogram

# Histogram for Fit across the sample
avg_Fit_dist <- ggplot(cbt_kff_joined_latent, aes(x = avg_Sess_Fit)) +
  geom_histogram(bins = 15,
    fill = "lightblue",
    color = "blue") +
  labs(title = "Distribution of Average Session Fit",
    x = "Average Session Fit",
    y = "Frequency") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
avg_Fit_dist

# Histogram for Tailoring
avg_tail_dist <- ggplot(cbt_kff_joined_latent, aes(x = avg_Sess_Tail)) +
  geom_histogram(bins = 15,
    fill = "lightgreen",
    color = "darkgreen") +
  labs(title = "Distribution of Average Session Tailoring",
    x = "Average Session Tailoring",
    y = "Frequency") +
  theme_classic(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
avg_tail_dist
```

```{r}

# Initial table of *descriptive statistics* with means, SDs, and lower triangle of the correlation matrix for the following variables: fit, tailoring, baseline anxiety severity, ACEs Total score, externalizing behavior problems, attention behavior problems, caregiver complexity, caregiver accommodation: overprotection, post-treatment anxiety severity, caregiver treatment satisfaction, youth treatment satisfaction


# Selecting the variables for the table and renaming them. This is subset or smaller dataset containing only the variables that we will need in the descriptive statistics / correlation table.

desc_data <- cbt_kff_joined_latent %>%
  select(
    `Fit` = avg_Sess_Fit,
    `Tailoring` = avg_Sess_Tail,
    `Baseline Anxiety Severity` = baseline_anxiety,
    `ACEs Total Score` = ACEQ_Tot_pool_B,
    `Externalizing Behavior Problems` = behavior_problems_ext,
    `Attention Behavior Problems` = behavior_problems_att,
    `Caregiver Complexity` = caregiver_complexity,
    `Caregiver Accommodation: Overprotection` = PABUover_C_B,
    `Post-treatment Anxiety Severity` = post_anxiety,
    `Caregiver Treatment Satisfaction` = SATQ_Mean_C_P_0,
    `Youth Treatment Satisfaction` = SATQ_Mean_Y_P_0)

# Computing descriptive statistics for each variable

# sapply() applies a function (like mean and sd) to each column of the subset dataset desc_data while na.rm = TRUE means missing values are ignored.
means <- sapply(desc_data, mean, na.rm = TRUE)
sds   <- sapply(desc_data, sd, na.rm = TRUE)

# Computing the correlation matrix: use = "pairwise.complete.obs" means each correlation is computed using all cases that have data on that specific pair of variables.
cor_mat <- cor(desc_data, use = "pairwise.complete.obs")

# Creating a display version of the correlation matrix where:
# - the diagonal is blank
# - the upper triangle is blank
# so only the lower triangle remains visible in the final table
cor_display <- cor_mat
cor_display[upper.tri(cor_display, diag = TRUE)] <- NA


# Now that we have both parts of the table, we can build the table data frame. This starts the table with:
# - Variable name
# - Mean
# - SD

table_df <- data.frame(Variable = names(desc_data),
  Mean = round(means, 2),
  SD = round(sds, 2),
  check.names = FALSE)
# the round(..., 2) command rounds values to 2 decimal places. The check.names = FALSE keeps column names exactly as written.


# Adding numbers to the variable names so the correlation columns can be matched to the numbered variable list.
table_df <- table_df %>%
  mutate(Variable = paste0(seq_along(Variable), ". ", Variable))

# Adding one correlation column at a time. The columns are named 1, 2, 3, etc. to match the numbered variable list.
for (i in seq_len(ncol(cor_display))) {table_df[[as.character(i)]] <- round(cor_display[, i], 2)}

# Replacing all NA values with blank cells for cleaner table display. This removes the visible NAs in the upper triangle and diagonal.
table_df[is.na(table_df)] <- ""


# Putting everything together and creating the formatted gt table
desc_gt <- table_df %>%
  gt() %>%
  # Converting the data frame into a gt table

  tab_header(title = md("**Descriptive Statistics and Correlations**")) %>%
  # Adding a title and bolding it using markdown syntax

  cols_label(Variable = "Variable",
    Mean = "M",
    SD = "SD") %>%
  # Renaming displayed column headers. Correlation columns keep their numbered labels automatically.

  opt_table_font(font = list("Times New Roman", "Times")) %>%
  # Setting the font to times new roman or times.

  cols_align(align = "left",
    columns = Variable) %>%
  # Left-aligning the variable names

  cols_align(align = "center",
    columns = -Variable) %>%
  # Center-aligning all other columns (M, SD, and correlations)

  cols_width(Variable ~ px(320),
    Mean ~ px(70),
    SD ~ px(70)) %>%
  # Setting column widths. This makes the variable-name column wide enough for long labels, while keeping Mean and SD narrow.

  tab_options(table.font.size = px(14),
    heading.title.font.size = px(16),
    data_row.padding = px(4),
    table.border.top.width = px(1.5),
    table.border.bottom.width = px(1.5),
    heading.border.bottom.width = px(1),
    table_body.border.bottom.width = px(1))
  # Controlling font size, title size, row spacing, and table borders to make the table look cleaner.

# Printing the final table
desc_gt

```


Second, it tests whether baseline child clinical characteristics, caregiver clinical complexity, and demographic factors predict these fit and tailoring indices. 

Model 1: Predicting Mean Fit
MeanFit_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i

```{r}

# Model 1 
model1 <- lm(avg_Sess_Fit ~ baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model1)

# The Main Outcomes paper controls: InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3
```

Model 2: Predicting Mean Tailoring
MeanTailoring_i=b_0+b_1 (BaselineAnxiety_i)+b_2 (BehaviorProblems_i)+b_3 (CaregiverComplexity_i)+b_4 (Age_i)+b_5 (Gender_i)+b_6 (CaregiverEducation_i)+e_i

```{r}

# Model 2
model2 <- lm(avg_Sess_Tail ~ baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model2)
```


```{r}
# Output table: Models 1 and 2

# -----------------------------
# Helper for APA-style p values
# -----------------------------
fmt_p <- function(p) {
  ifelse(p < .001, "< .001", sprintf("%.3f", p))
}

# -----------------------------
# Clean predictor labels
# -----------------------------
term_labels <- c(
  "(Intercept)" = "Intercept",
  "baseline_anxiety" = "Baseline anxiety severity",
  "behavior_problems_ext" = "Externalizing behavior problems",
  "behavior_problems_att" = "Attention behavior problems",
  "ACEQ_Tot_pool_B" = "ACEs total score",
  "caregiver_complexity" = "Caregiver complexity",
  "PABUover_C_B" = "Caregiver accommodation",
  "Educ3_C" = "Caregiver education",
  "InsecurityStatus_Pool" = "Resource instability",
  "Male_Y_pool" = "Male",
  "AgeGrp_0" = "Age group",
  "Lang_0" = "Spanish language",
  "Black_Y_pool" = "Black",
  "Hisp_Y_pool" = "Hispanic",
  "Site1" = "Baltimore",
  "Site2" = "Boston",
  "Site3" = "Miami"
)

# -----------------------------
# Tidy each model
# -----------------------------
model1_tidy <- tidy(model1) %>%
  mutate(
    term = recode(term, !!!term_labels),
    Estimate_1 = round(estimate, 2),
    SE_1 = round(std.error, 2),
    t_1 = round(statistic, 2),
    p_1 = fmt_p(p.value)
  ) %>%
  select(term, Estimate_1, SE_1, t_1, p_1)

model2_tidy <- tidy(model2) %>%
  mutate(
    term = recode(term, !!!term_labels),
    Estimate_2 = round(estimate, 2),
    SE_2 = round(std.error, 2),
    t_2 = round(statistic, 2),
    p_2 = fmt_p(p.value)
  ) %>%
  select(term, Estimate_2, SE_2, t_2, p_2)

# -----------------------------
# Model-level fit statistics
# -----------------------------
model1_glance <- glance(model1)
model2_glance <- glance(model2)

fit_rows <- tibble(
  term = c("R²", "Adjusted R²", "F statistic"),
  Estimate_1 = c(
    round(model1_glance$r.squared, 2),
    round(model1_glance$adj.r.squared, 2),
    round(model1_glance$statistic, 2)
  ),
  SE_1 = c(NA_real_, NA_real_, NA_real_),
  t_1 = c(NA_real_, NA_real_, NA_real_),
  p_1 = c(NA_character_, NA_character_, fmt_p(model1_glance$p.value)),
  Estimate_2 = c(
    round(model2_glance$r.squared, 2),
    round(model2_glance$adj.r.squared, 2),
    round(model2_glance$statistic, 2)
  ),
  SE_2 = c(NA_real_, NA_real_, NA_real_),
  t_2 = c(NA_real_, NA_real_, NA_real_),
  p_2 = c(NA_character_, NA_character_, fmt_p(model2_glance$p.value))
)

# -----------------------------
# Combine everything
# -----------------------------
results_table <- full_join(model1_tidy, model2_tidy, by = "term") %>%
  bind_rows(fit_rows)

# Put rows in the order you want
row_order <- c(
  "Intercept",
  "Baseline anxiety severity",
  "Externalizing behavior problems",
  "Attention behavior problems",
  "ACEs total score",
  "Caregiver complexity",
  "Caregiver accommodation",
  "Caregiver education",
  "Resource instability",
  "Male",
  "Age group",
  "Spanish language",
  "Black",
  "Hispanic",
  "Baltimore",
  "Boston",
  "Miami",
  "R²",
  "Adjusted R²",
  "F statistic"
)

results_table <- results_table %>%
  mutate(term = factor(term, levels = row_order)) %>%
  arrange(term) %>%
  mutate(term = as.character(term))

# -----------------------------
# Make gt table
# -----------------------------
reg_gt <- results_table %>%
  gt(rowname_col = "term") %>%
  tab_header(
    title = md("**Table 2. Main Model Results**")
  ) %>%
  tab_spanner(
    label = "Model 1: Predicting Mean Fit",
    columns = c(Estimate_1, SE_1, t_1, p_1)
  ) %>%
  tab_spanner(
    label = "Model 2: Predicting Mean Tailoring",
    columns = c(Estimate_2, SE_2, t_2, p_2)
  ) %>%
  cols_label(
    Estimate_1 = "Estimate",
    SE_1 = "SE",
    t_1 = "t(92)",
    p_1 = md("*p*"),
    Estimate_2 = "Estimate",
    SE_2 = "SE",
    t_2 = "t(92)",
    p_2 = md("*p*")
  ) %>%
  fmt_number(
    columns = c(Estimate_1, SE_1, t_1, Estimate_2, SE_2, t_2),
    decimals = 2
  ) %>%
  fmt_missing(
    columns = everything(),
    missing_text = ""
  ) %>%
  opt_table_font(
    font = list(
      "Times New Roman",
      "Times",
      "Liberation Serif",
      "Nimbus Roman",
      "serif"
    )
  ) %>%
  cols_align(
    align = "left",
    columns = stub()
  ) %>%
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  cols_width(
    stub() ~ px(260),
    everything() ~ px(90)
  ) %>%
  tab_source_note(
    source_note = md(
      "*Note.* Entries are unstandardized regression coefficients, standard errors, t statistics, and exact p values. R² = proportion of variance explained by the model. Both models were estimated with listwise deletion in `lm()`. Model 1 outcome = average session fit; Model 2 outcome = average session tailoring."
    )
  ) %>%
  tab_options(
    table.font.size = px(13),
    heading.title.font.size = px(15),
    data_row.padding = px(4),
    table.border.top.width = px(1.5),
    table.border.bottom.width = px(1.5),
    heading.border.bottom.width = px(1),
    table_body.border.bottom.width = px(1)
  )

reg_gt

```


Third, it tests whether average fit and average tailoring predict anxiety response outcome and treatment satisfaction (for caregiver and child). 

Model 3: Predicting Anxiety Outcome
PostAnxiety_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

```{r}

# Model 3: includes both flexibility variables
model3 <- lm(post_anxiety ~ avg_Sess_Fit + avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model3)

# Model 3a: includes only fit
model3a <- lm(post_anxiety ~ avg_Sess_Fit + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model3a)

# Model 3b: includes only tailoring
model3b <- lm(post_anxiety ~ avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model3b)

# Model 3 Quadratic: with only tailoring
model3b_quad <- lm(post_anxiety ~ avg_Sess_Tail + I(avg_Sess_Tail^2) + baseline_anxiety + behavior_problems_ext + behavior_problems_att + ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, data = cbt_kff_joined_latent)

summary(model3b_quad)

```


Model 4: Predicting Caregiver Treatment Satisfaction
CaregiverSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

```{r}
model4 <- lm(SATQ_Mean_C_P_0 ~ avg_Sess_Fit + avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model4)

model4a <- lm(SATQ_Mean_C_P_0 ~ avg_Sess_Fit + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model4a)

model4b <- lm(SATQ_Mean_C_P_0 ~ avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model4b)
```


Model 5: Predicting Youth Treatment Satisfaction
YouthSatisfaction_i=b_0+b_1 (MeanFit_i)+b_2 (MeanTailoring_i)+b_3 (BaselineAnxiety_i)+b_4 (BehaviorProblems_i)+b_5 (CaregiverComplexity_i)+b_6 (Age_i)+b_7 (Gender_i)+b_8 (CaregiverEducation_i)+e_i

```{r}
model5 <- lm(SATQ_Mean_Y_P_0 ~ avg_Sess_Fit + avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model5)

model5a <- lm(SATQ_Mean_Y_P_0 ~ avg_Sess_Fit + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model5a)

model5b <- lm(SATQ_Mean_Y_P_0 ~ avg_Sess_Tail + baseline_anxiety + behavior_problems_ext + behavior_problems_att +  ACEQ_Tot_pool_B + caregiver_complexity + PABUover_C_B + Educ3_C + InsecurityStatus_Pool + Male_Y_pool + AgeGrp_0 + Lang_0 + Black_Y_pool + Hisp_Y_pool + Site1 + Site2 + Site3, cbt_kff_joined_latent)
summary(model5b)
```

